% LaTeX file for resume 
% This file uses the resume document class (res.cls)

\documentclass[margin,line]{res} 
% the margin option causes section titles to appear to the left of body text 
\textwidth=5.2in % increase textwidth to get smaller right margin
%\usepackage{helvetica} % uses helvetica postscript font (download helvetica.sty)
%\usepackage{newcent}   % uses new century schoolbook postscript font 
\usepackage{hyperref}

\newcommand{\code}[1]{\,{\tt #1}\,}
\newcommand{\C}[1]{\code{#1}}

\begin{document} 
 
\name{Gowtham Kaki} 
 
%\fontsize{9pt}{1} \selectfont
\address{https://gowthamk.github.io \hspace{1.35in} gkaki@purdue.edu
          \hspace{1.25in} (201) 417 1775 }

 
\begin{resume} 
 
\section{Research Focus}
Applying language design and automated reasoning techniques to
  simplify programming complex real-world systems.

\section{Education} 
{\bf Ph.D., Computer Science}. \hfill 2012 - present\\
Advised by Prof. Suresh Jagannathan.\\
Purdue University, West Lafayette, IN.\\
\\
{\bf M.S., Computer Science}. \hfill May, 2016\\
Purdue University, West Lafayette, IN.\\
GPA: 3.77/4.0 \\
\\
{\bf B.E.(Hons)., Computer Science}. \hfill July, 2009\\
BITS, Pilani, India.\\
GPA: 8.68/10.0
 

\section{Publications}

Gowtham Kaki, KC Sivaramakrishnan, and Suresh Jagannathan. Version
  Control is for Your Data Too. In the proceedings of Symposium on
  Advances in Programming Languages (SNAPL), 2019.
\vspace*{0.1in}\\
Gowtham Kaki, Kapil Earanky, KC Sivaramakrishnan, and Suresh
  Jagannathan. Safe Replication through Bounded Concurrency
  Verification. In the proceedings of ACM SIGPLAN Conference on
  Object-Oriented Programming Systems Languages and Applications
  (OOPSLA), 2018.
\vspace*{0.1in}\\
Gowtham Kaki, and G Ramalingam. Safe Transferable Regions.  In the
  proceedings of the European Conference on Object-Oriented
  Programming (ECOOP), 2018.
\vspace*{0.1in}\\
Gowtham Kaki, Kartik Nagar, Mahsa Najafzadeh, and Suresh Jagannathan.
  Alone Together: Compositional Reasoning and Inference for Weak
  Isolation. In the proceedings of ACM SIGPLAN Conference on
  Principles of Programming Langauges (POPL), 2018.
\vspace*{0.1in}\\
KC Sivaramakrishnan, Gowtham Kaki, and Suresh Jagannathan.
  Representation without Taxation: A Uniform, Low-Overhead, and
  High-Level Interface to Eventually Consistent Key-Value Stores. In
  IEEE Data Engineering Bulletin, 2016.
\vspace*{0.1in}\\
KC Sivaramakrishnan, Gowtham Kaki, and Suresh Jagannathan.
  Declarative Programming over Eventually Consistent Data Stores. In
  the proceedings of ACM SIGPLAN Conference on Programming Languages
  Design and Implementation (PLDI), 2015.
\vspace*{0.1in}\\
Gowtham Kaki, and Suresh Jagannathan. A Relational Framework for
  Higher-Order Shape Analysis. In the proceedings of ACM SIGPLAN
  International Conference on Functional Programming (ICFP), 2014.
\vspace*{0.1in}\\
Sunita Bansal, Gowtham Kaki, and Chittaranjan Hota. A Novel Adaptive
  Scheduling Algorithm for Computational Grids. In the proceedings of
  IEEE International Conference on Internet Multimedia Systems
  Architecture and Application (IMSAA), 2009.

\section{Preprints/Drafts}
Gowtham Kaki, Swarn Priya, KC Sivaramakrishnan, and Suresh
  Jagannathan. Relational Reasoning for Mergeable Replicated Data
  Types. Under submission to OOPSLA 2019.
\vspace*{0.1in}\\
Gowtham Kaki, KC Sivaramakrishnan, Samodya Abeysiriwardane, and Suresh
  Jagannathan. Mergeable Types. Extended abstract presented at ML
  Workshop, 2017.
\vspace*{0.1in}\\
Gowtham Kaki, and Suresh Jagannathan. A Lightweight Symbolic Execution
Framework for Ruby-on-Rails. Extended abstract presented at PLDI 2016
Student Research Competition (SRC).
% \vspace*{0.1in}\\
% Gowtham Kaki, Kapil Earanky, KC Sivaramakrishnan, and Suresh
% Jagannathan. From {\sf ACID} to {\sf BASE}: Effective Verification of
% Highly-Scalable Transactions through Bounded Staleness. Draft
% available on request.
 
\section{Professional \\ Experience}

{\bf Research Intern, Microsoft Research India, Bangalore} \\
(May - August, 2014 \& July-August, 2015) 
Built a region type system and region type inference to ensure the
safety of dataflow programs that rely on programmer-managed memory
regions, instead of garbage collection for memory management.

{\bf Software Engineer, Yahoo SDC, Bangalore, India} \hfill August, 2009 -
July, 2011\\
Frontend engineering for Yahoo content platforms group. Developed AJAX
and php tools for querying, processing and presenting
loosely-structured data from various content grids inside yahoo. The
tools were used by Yahoo's content curators.

{\bf Engineering intern, Qualcomm, Hyderabad, India} \hfill January - June, 2009\\
QA Engineering for Application-specific integrated circuit (ASIC) -
User interface module (UIM) group. Developed tools to test low-level
mobile network code.

% \section{Course Work}
% Design \& Analysis of Algorithms (A), Programming Languages (A),
% Software Engineering (B), Metaprogramming and Program Generation (A+),
% Distributed Systems (A), Parallel Computing (A), and Current topics in
% Theoretical Computer Science (A-).  

% \section{Teaching Assistantship}
% \begin{itemize}
% \item CS240 C Programming. Fall 2012.
% \item CS565 Programming Languages. Spring, 2013.
% \end{itemize}

\section{Awards and Fellowships}
\begin{itemize}
  \item BITS Alumni Association (BITSAA)'s ``30 under 30''
    award for excellence in academic research\footnote{\url{https://www.bitsaa.org/page/30U30}}.
  \item Purdue CS's Maurice H. Halstead Memorial Award (2019) for
    outstanding research in Software Engineering\footnote{\url{https://www.cs.purdue.edu/news/articles/2019/2019-cs-award-banquet.html}}.
  \item Google PhD Fellowship, 2018\footnote{\url{https://ai.googleblog.com/2018/04/announcing-2018-google-phd-fellows-for.html}}.
% \item Invited to Dagstuhl seminar on Language-based Verification
%   Tools (March, 2015).
% \item Invited to Shonan semninar on Programming Languages for
%   Data-Intensive Applications (July, 2019).
  \item ACM SIGPLAN PAC funding for ICFP'14 and OOPSLA'18.
  \item ECOOP travel funding, 2018.
  \item ACM SIGPLAN PLMW scholarship, 2014.
  \item Institute merit-cum-need scholarship during all semesters of
    my undergraduate education at BITS Pilani, India.
\end{itemize}

\section {Professional Service}
\begin{itemize}
  \item Served on Artifact Evaluation Committees (AECs) of PLDI'16,
    ICFP'17, ICFP'18, ECOOP'19, and PLDI'19.
  \item Started the Purdue PL (PurPL) reading group in Fall, 2013,
    which has since taken a life of its own to become an umbrella
    organization for all PL groups at Purdue\footnote{
    \url{http://purduepl.github.io/}}.
  \item Served on Purdue CS Graduate Student Board (GSB), 2011-2012.
  \item Served as the Secretary of Computer Science Association (CSA)
    at BITS Pilani, 2007-08. 
\end{itemize}

\section{References}
Prof. Suresh Jagannathan (Purdue University),
Dr. Ganesan Ramalingam (Microsoft Research), \&
Prof. KC Sivaramakrishnan (IIT Madras).
Other references will be available on request.
\end{resume} 
\end{document} 



