This diversity statement is in response to the following question:
\emph{Please give an example of when you have changed or altered your
approach to research based on your knowledge and awareness of
underserved or underrepresented populations, and/or underreporting of
relevant data stemming from explicit or implicit bias.}

I have long been mindful of the barriers faced by the underrepresented
groups in academia and tech industry. My own experiences of growing up
in an economically- and socially-backward community in rural India,
and later studying at an acclaimed private university with an
overwhelmingly upper-class student population has made me understand
the pernicious effects of exclusion, real or perceived, on one's
academic performance and confidence. 

In my capacity as a course instructor at Purdue this semester, I have
taken steps to ensure that every student feels welcome in the
classroom, and included in the classroom discussion. Among the things
I undertook is an effort to address the gender stereotype threat
facing women in CS by carefully choosing examples, case studies and
role models that \emph{prime} the students with positive
stereotypes~\cite{MA07,SG11}.  Another conscious effort was aimed at
disadvantaged students whose education has suffered due to the lack of
resources and unfavorable social conditions in inner cities and
developing world. It is unfair to expect such students to have the
same level of confidence as the privileged students, and therefore
unreasonable to assume that they will participate in classroom
discussion, or take advantage of learning resources without any
external intervention. As an instructor, I have always strived to
include disadvantaged students in the classroom discussion by
fostering a conducive environment where everyone is encouraged to
express their opinion without the fear of judgment, and every opinion
is given a thoughtful consideration. I have also carefully refrained
from speech or action that could directly or indirectly reinforce
informal social hierarchies among students that contribute to
defensive social climate~\cite{BGJ02}. 

I have also long been committed to equity, diversity, and inclusion in
my research, mentorship and professional activities. Over the course
of years, I have developed a positive approach towards research
collaboration and mentorship that emphasizes strengths of my
collaborators while acknowledging and embracing their weaknesses. This
approach has helped me build great professional relationships with my
fellow researchers, particularly my female colleagues. Among my
mentees is a freshman female graduate student, who is now a productive
researcher, and my immediate co-author on a recently published journal
paper. Beyond the formal channels, I frequently engage my colleagues
in conversations about diversity and inclusion, which I found to be
effective in questioning stereotypes, raising awareness about implicit
biases, and understanding cultural differences. As an assistant
professor in the future, I plan to undertake more initiatives aimed at
fostering a sense of belongingness in grad school among students from
vulnerable groups. In particular, I am committed to the creation of a
forum for Women in Software Engineering (WiSE), which understands and
addresses the issues faced by women in software engineering research,
and works towards making the research community more inclusive. I
would also like to constitute a forum to discuss mental health issues
in academia, which seem to disproportionately affect students from
underrepresented groups.  Furthermore, I will continue engaging the
academic community in conversations about empathy, equality, diversity
and inclusion, which I believe are necessary to substantiate the
institutional efforts towards furthering these causes.

